\mytitle

### 2016-04-25 report/paper.tex
Correct errors

Distributed systems, like centralized systems needs to preserve the temporal order of event produced by concurrent process in the system. When there are spatially separated processes that can only communicate through messages, you can’t tell what in the system happened before something else.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of order \autocite{lamport1978time}:
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

@@

\section{EpTO}
\todo[inline]{Here we will present EpTO and the different components}
\begin{figure}
	\includegraphics[width=\linewidth]{figures/epto-architecture.png}
	\caption{\todo[inline]{We need a better quality figure (SVG)}\protect\footnotemark{}.}
	\label{fig:epto-architecture}
\end{figure}
\subsection{Definitions}
A process or peer is defined as an actor in our system running the application that needs total order. Each process will communicate with other processes in the distributed systems and exchange events and order them together.
\par
An event is defined as something that happens at a given time. For example, we could imagine a system where each process publish some sort of data. The moment where this data is published combined with the data is called an event.
\par
A ball is a set of events bundled together and sent as one package. We use balls in EpTO to reduce network traffic and make it scalable in terms of processes and events.

@@

\todo[inline]{Maybe speak about holes created by the probabilistic nature of the algorithm}
\par
Since EpTO is using a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability). In this case there will be a hole in the sequence of delivered events but even in this case, the order of the delivered event will be protected by EpTO's deterministic ordering algorithm and total order property is preserved.
\subsection{Dissemination Component}
The Dissemination component is the component that bridges EpTO with the rest of the network. As we can see in Figure \ref*{fig:epto-architecture}. It is in charge of receiving balls, open them , pass them to the Ordering component and then forward them to \textit{K} other processes, where K denotes the fan-out of the gossip.
\par
When an application wants to publish an event, it will broadcast this event to the Dissemination component.

\subsection{Ordering Component}
The Ordering component is responsible for orders events before delivering them to the application.
To do so, this component has a \textit{received} hashmap of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hashmap containing all the events which are delivered.
\par
In brief, ordering component first increment the timestamp of all the events which has been received in previous rounds to indicate the start of a new round. Then it process new events in the received ball by discarding events that have been received already (delivered events or events with timestamp smaller than the last delivered event) to prevent delivering duplicate event, remaining event in the received ball will add to the \textit{received} hashmap, and wait to be delivered based on the stability oracle and other undelivered events on each round.
\subsection{Stability Oracle}
The Stability oracle is the component that outputs timestamps. It will increment its local clock every round as well as synchronize itself using timestamps of newly received events, to make sure the local clock does not drift too much.
\par
This Stability oracle offers an API to the Ordering component letting it know when an event is mature enough to be delivered to the application.
\par
As this component is local and only correct itself when we receive new events, it generates no network traffic. This means it doesn't impact the scalability of EpTO.
\section{Performance Comparisons}
\todo[inline]{Here we will present the different test we ran, the methodology used and the result we obtained}
\subsection{Methodology}
\todo[inline]{Here we explain how we ran our tests and on what type of machines}
\subsection{Peer membership known}
\todo[inline]{We will present the results obtained and try to explain the results}
\subsection{Peer membership unknown}
\todo[inline]{We will present the results obtained and try to explain the results}

@@

% use section* for acknowledgement
\section*{Acknowledgment}

>>>>

Distributed systems, like centralized systems need to preserve the temporal order of event produced by concurrent process in the system. When there are separated processes that can only communicate through messages, you can’t easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}, the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

@@

\section{EpTO}
\begin{figure}
	\includegraphics[width=\linewidth]{figures/epto-architecture.png}
	\caption{\todo[inline]{We need a better quality figure (SVG)}\protect\footnotemark{}.}
	\label{fig:epto-architecture}
\end{figure}
\subsection{Definitions}
A process or peer is defined as an actor in our system running the application that needs total order. Each process will communicate with other processes in the distributed systems and exchange events and order them together.
\par
An event is defined as something that happens at a given time. For example, we could imagine a system where each process publish some sort of data. The moment where this data is published combined with the data is called an event.
\par
A ball is a set of events bundled together and sent as one package. We use balls in EpTO to reduce network traffic and make it scalable in terms of processes and events.

@@

\par
Since EpTO is using a probabilistic agreement instead of a deterministic agreement, there might be a situation where a peer does not receive an event (with a very low arbitrary probability). In this case there will be a hole in the sequence of delivered events but even in this case, the order of the delivered event will be protected by EpTO's deterministic ordering algorithm and total order property is preserved.
\subsection{Dissemination Component}
The Dissemination component is the component that bridges EpTO with the rest of the network. As we can see in Figure \ref*{fig:epto-architecture}, it is in charge of receiving balls, open them , pass them to the Ordering component and then forward them to \textit{K} other processes, where K denotes the gossip fan-out.
\par
When an application wants to publish an event, it will broadcast this event to the Dissemination component.

\subsection{Ordering Component}
The Ordering component is responsible for ordering events before delivering them to the application.
To achieve this, the Ordering component has a \textit{received} hash table of (\textit{id}, \textit{event}) pairs containing all the events which are received by the peer, but not yet delivered to the application and a \textit{delivered} hash table containing all the events which are delivered.
\par
In brief, the Ordering component first increments the timestamp of all the events which have been received in previous rounds to indicate the start of a new round. Then, it processes new events in the received ball by discarding events that have been received already (delivered events or events with timestamp smaller than the last delivered event). This is done to prevent delivering duplicate events. The remaining events in the received ball will be added to the \textit{received} hash table, and wait to be delivered based on the Stability oracle.
\subsection{Stability Oracle}
The Stability oracle is the component that outputs timestamps. It will increment its local clock every round as well as synchronize itself using timestamps of newly received events, to make sure the local clock does not drift too much.
\par
This Stability oracle offers an API to the Ordering component letting it know when an event is mature enough to be delivered to the application.
\par
As this component is local and only correct itself when we receive new events, it generates no network traffic. This means it does not impact the scalability of EpTO.
\section{Performance Comparisons}
\todo[inline]{Here we will present the different test we ran, the methodology used and the result we obtained}
\subsection{Methodology}
\todo[inline]{Here we explain how we ran our tests and on what type of machines}
\subsection{Peer membership known}
\todo[inline]{We will present the results obtained and try to explain the results}
\subsection{Peer membership unknown}
\todo[inline]{We will present the results obtained and try to explain the results}

@@

% use section* for acknowledgement
\section*{Acknowledgment}
\todo[inline]{Write acknowledgment}
